Preprocessing datasets

### 2020-05-14 datasets/readme.Rmd
update dataset preprocessing'

---
title: "Dataset preprocessing"
output: github_document
---
This notebook processes all systematic review datasets into datasets suitable testing ASReview.

Every dataset is a .csv file with the following columns: 
  - title
  - abstract
  - label_included
  - keywords? (only for hyperparameter optimization, not for simulation).

Entries with missing abstracts are removed

@@

### Set-up
```{r, echo = FALSE}
datasets <- c("ace", "nudging", "ptsd", "software", "virus", "wilson")
nd <- length(datasets)
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

@@

## ACE
```{r}
ace <- template
# paper ------------------------------------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

@@

# asreview ---------------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)
# replace empty abstracts by NA values.

ace_asr[ace_asr$abstract == "", "abstract"] <- NA
# stats on missingness in this data
drops["ace",] <- ace_asr %>%
           summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

ace["asreview", ] <- c(nrow(ace_asr),
                       NA,
                       sum(ace_asr$label_included))
```

### Create test dataset
```{r}
# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

@@

### Some statistics 
```{r}
# missingness and duplicates in raw data on ASReview
drops["ace",]
ace
```

## nudging dataset 
```{r}
# by Rosanna Nagtegaal
nudging <- template
nudging["paper", ] <- c(2006, 377, 100) 
```

@@

### Create test dataset

### Some statistics
```{r}

drops["nudging",]
nudging
```

@@

Derive statistics in final test set.
```{r}
drops["ptsd", ]
ptsd
```

@@

### Statistics 
```{r}
drops["software", ] <- 
  hall_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

hall
```

@@

@@

Statistics
```{r}
drops["wilson",]

wilson
```

@@

# save datafile, to serve as data for descriptive table in manuscript
saveRDS(all, file = "data_statistics/all.RDS")

saveRDS(drops, file = "data_statistics/drops.RDS")
```

@@

       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       `fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")) %>% 
  t() %>% 
  kable("markdown") 
# %>%
#   kable_styling(full_width = TRUE) %>%
#   #scroll_box(width = "100%", height = "200px") %>%
#   pack_rows("Paper", 2, 5) %>%
#   pack_rows("Raw data", 6, 9) %>%
#   pack_rows("ASReview data", 10, 13) %>%
#   pack_rows("Test data set", 14, 17)  
```

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("markdown",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2) 
# %>%
#   kable_styling(full_width = TRUE)

@@

>>>>

---
title: "Preprocessing datasets"
output: github_document
bibliography: ../manuscript/manuscript/asreview.bib
---

This notebook processes six systematic review datasets into datasets suitable for simulating the screening process.

- __Wilson__ - is on a review on effectiveness and safety of treatments of Wilson Disease, a rare genetic disorder of copper metabolism. Dataset: [@Appenzeller-Herzog2020]. Paper: [@Appenzeller-Herzog2019]. 
- __Ace__ -  contains publications on the efficacy of Angiotensin-converting enzyme (ACE) inhibitors, a drug treatment for heart disease Paper and dataset: [@Cohen2006]. 
- __Virus__ -  is from a systematic review on studies that performed viral Metagenomic Next-Generation Sequencing (mNGS) in farm animals [@Kwok2020]. 
- __Software__ - from the software engineering field, contains publications on fault prediction in source code [@Hall2012].
- __Nudging__ dataset - [@Nagtegaal2019a] on nudging healthcare professionals [@Nagtegaal2019], stemming from the area of 
- __PTSD__ - on studies applying latent trajectory analyses on posttraumatic stress after exposure to traumatic events [@vandeSchoot2017].

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data.

Every dataset has 2 versions: one for hyperparameter optimization (`test_datasets`), and one for simulation (`sim_datasets`).

Every simulation dataset is a .csv file with the columns `title`, `abstract`, `label_included`, the last one indicating whether a publication was relevant or not. This is the label that is queried by the active learning model when simulating a systematic review. The reason for this sparse datasets (no keywords are included etc) is twofold. 1) This is by PRISMA guidelines the only information that should be used to identify relevant publications from a search and 2) I wanted equal information for all datasets, and for some datasets this was the only information that was available. For hyperparameter optimization I used all information accessible to arrive at better hyperparameters, with future projects in mind.

@@

### Preprocessing datasets
For every datasets, 3 numbers are given: total number of publications, abstract inclusions and final inclusions. For some datasets there are discrepancies between how these numbers are reported in the publication and how they actually are in the raw dataset. Therefore, these numbers are given for how they are reported in the manuscript, how they are found in the raw data, and how they are found in the test dataset, e.g. after removing duplicates and missing abstracts. Moreover, statistics on missingness and duplicates are given.

```{r, echo = FALSE}
datasets <- c("ace", "nudging", "ptsd", "software", "virus", "wilson")
nd <- length(datasets)
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

@@

## Ace dataset
```{r}
ace <- template

# information given in the paper -----------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

@@

# asreview dataset -------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)
```

### Create test dataset
```{r}
# replace empty abstracts by NA values. 
ace_asr[ace_asr$abstract == "", "abstract"] <- NA
# stats on missingness in this data
drops["ace",] <- ace_asr %>%
           summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

ace["asreview", ] <- c(nrow(ace_asr),
                       NA,
                       sum(ace_asr$label_included))

# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

@@

### Some statistics 
On missingness (# and %) and duplicates (# and %): 
```{r, echo = false}
# missingness and duplicates in raw data on ASReview
drops["ace",] %>% knitr::kable(format = "markdown", digits = 2)
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
ace %>% knitr::kable(format = "markdown")
```

## Nudging dataset 
```{r}
# by Rosanna Nagtegaal
nudging <- template
nudging["paper", ] <- c(2006, 377, 100) 
```

@@

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["nudging",]%>% knitr::kable(format = "markdown")
```
On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
nudging %>% knitr::kable(format = "markdown")
```

@@

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["ptsd",]%>% knitr::kable(format = "markdown")
```
On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
ptsd %>% knitr::kable(format = "markdown")
```

@@

```{r}
drops["software", ] <- 
  hall_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["hall",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
hall %>% knitr::kable(format = "markdown")
```

@@

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["virus",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
virus %>% knitr::kable(format = "markdown")
```

@@

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["wilson",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
wilson %>% knitr::kable(format = "markdown")
```

@@

# save datafile, to serve as data for descriptive table in manuscript
saveRDS(all, file = "data_statistics/all.RDS")
saveRDS(drops, file = "data_statistics/drops.RDS")
```

@@

       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       `fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")) %>% 
  t() %>% 
  kable("markdown", digits = 2) 
# %>%
#   kable_styling(full_width = TRUE) %>%
#   #scroll_box(width = "100%", height = "200px") %>%
#   pack_rows("Paper", 2, 5) %>%
#   pack_rows("Raw data", 6, 9) %>%
#   pack_rows("ASReview data", 10, 13) %>%
#   pack_rows("Test data set", 14, 17)  
```

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("markdown",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2)

@@

```
# refrences
